---
title: 'Recreating plots in R: the power of tiny gains'
author: Christopher Yee
date: '2021-09-21'
slug: recreating-plots-in-r-the-power-of-tiny-gains
categories:
  - Data Visualization
editor_options: 
  chunk_output_type: console
---

Once in awhile I find some great charts which make me stop and think.

Below is one example of this:

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">1% better every day is all it takes to completely change your life. <a href="https://t.co/kYbgGOzzZv">pic.twitter.com/kYbgGOzzZv</a></p>&mdash; Santiago (@svpino) <a href="https://twitter.com/svpino/status/1436404061985812483?ref_src=twsrc%5Etfw">September 10, 2021</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

I really appreciate the inspirational message behind it: being deliberate about the change you wish to see in the world.

With that in mind, I want to recreate that chart below using the [R programming language]((https://twitter.com/search?q=%23rstats)) (regardless if the trajectory path is realistic or not).


## Create data
```{r message=FALSE, warning=FALSE}
library(tidyverse)

df <- seq(1, 365, 1) %>% 
  as_tibble() %>% 
  rename(time = value) %>% 
  mutate(
    
    # 1% better every day
    better = 1.01^time,
    
    # 1% worse every day
    worse = 0.99^time
    
    ) %>% 
  pivot_longer(cols = 2:3)
```

## Visualize data
```{r message=FALSE, warning=FALSE}
cpalette <- c("seagreen", "salmon")

df %>% 
  ggplot(aes(time, value, color = name)) +
  geom_line(size = 1, arrow=arrow()) +
  
  # dotted line at 1
  geom_hline(yintercept = 1, lty = 2) +
  
  # side and bottom grid lines
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = -2) +
  
  labs(x = "Time", 
       y = str_wrap("Improvement or Decline", width = 10),
       title = "The Power of Tiny Gains") +
  scale_color_manual(values = cpalette) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_text(angle = 0, vjust = 0.5),
        legend.position = 'none')
```

<br>


## Related

* [Recreating plots in R: intro to bootstrapping](https://www.christopheryee.org/blog/recreating-plots-in-r-intro-to-bootstrapping/)

<br>


