---
title: '[Updated] Top Industries from Inc.5000 Companies'
author: Christopher Yee
date: '2020-01-07'
slug: top-industries-from-inc-5000-fastest-growing-companies-in-2019
categories:
  - Data Visualization
---

## Changelog

* Originally published on _September 10th, 2019_
* Built a [Shiny app for this](https://eeysirhc.shinyapps.io/inc5000_companies/)
* Full code can be found on [GitHub](https://github.com/Eeysirhc/inc5000_companies)

<hr>

One of my favorite online marketers, (the) [Glen Allsopp](https://twitter.com/ViperChill), tweeted the following:

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Over the past few weeks I&#39;ve went through every site in the Inc. 5000. My mind has been blown multiple times. Don&#39;t click if you&#39;re easily distracted. Enjoy! <a href="https://t.co/mHVK8rvb9X">https://t.co/mHVK8rvb9X</a> <a href="https://t.co/BoEb3qQ7LZ">pic.twitter.com/BoEb3qQ7LZ</a></p>&mdash; Glen Allsopp (@ViperChill) <a href="https://twitter.com/ViperChill/status/1166324614877679617?ref_src=twsrc%5Etfw">August 27, 2019</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

The public spreadsheet contains four fields:

* Inc.com URL
* URLs (company website)
* Revenue
* 3-Year-Growth

Although helpful I thought it would be interesting to explore the additional variables found on each company's INC profile page.

Thus, I fired up R and scaped the list of URLs to answer the question: *which industries were surveyed the most and how much revenue did they generate in 2019?*

## tl;dr

```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
##### HIDE #####
library(tidyverse) 

company_data <- read_csv("~/Development/inc5000_companies/companies_data_final.csv")

company_parsed <- company_data %>% 
  group_by(industry) %>% 
  summarize(revenue_millions = sum(revenue_millions),
            count = n()) %>% 
  ungroup() %>% 
  filter(!is.na(revenue_millions)) %>% 
  mutate(pct_count = count / sum(count),
         pct_revenue = revenue_millions / sum(revenue_millions),
         revenue_billions = revenue_millions / 1000) 

top_20 <- company_parsed %>% 
  arrange(desc(pct_revenue)) %>% 
  select(industry, revenue_billions, count, pct_revenue, pct_count) %>% 
  top_n(20)
##### HIDE #####
```


```{r message=FALSE, warning=FALSE, echo=FALSE}
library(plotly)

plot_ly(data = company_parsed, 
        x = ~pct_count, 
        y = ~pct_revenue, 
        mode = "scatter",
        type = "scatter",
        size = 10,
        color = ~industry,
        colors = 'Set1', 
        hoverinfo = "text",
        text = ~paste("<b>Industry:</b> ",industry, 
                     "<br><b>Total Companies:</b> ", count,
                      "<br><b>Total Revenue ($Bn):</b> ", revenue_billions),
        showlegend = FALSE) %>% 
  layout(xaxis = list(title = "% of Total Companies",
                      tickformat = "%"),
         yaxis = list(title = "% of Total Revenue",
                      tickformat = "%"))
```


## Load packages
```{r eval=FALSE}
library(tidyverse)
library(rvest) # SIMPLE WEB SCRAPING
```

## Get URLs from CSV
```{r eval=FALSE}
inc5000 <- read_csv("inc5000_fastest_growing_companies.csv") %>% 
  rename(urls = `Inc.com URL`,
         website = URLs)
```

## Base R format
We temporarily need to move out of the tidyverse and leverage base R for the next step.
```{r eval=FALSE}
company_urls <- inc5000$urls 
```

## Loop function
A for loop is required to gather our data with the following order of operations:

* Take URL from list
* Crawl the page
* Extract the page elements we want
* Store into data frame
* Rinse and repeat from step 1

```{r eval=FALSE}
# INITIALIZE DATA FRAME
company_raw <- data.frame()

for (page_url in company_urls){
  print(page_url)
  
  # RETRIEVE INC5000 PROFILE PAGE
  page <- read_html(page_url)
  
  # PARSE REVENUE
  revenue_millions <- page %>% 
    html_nodes(xpath = '//*/section[1]/div[3]/dl[1]/dd') %>% 
    html_text() %>% 
    str_replace(" Million", "") %>% # STRIP 'MILLIONS' FROM DATA VALUE
    str_replace("\\$", "") # STRIP $ SIGN FROM DATA VALUE

  # PARSE INDUSTRY
  industry <- page %>% 
    html_nodes(xpath = '//*/section[1]/div[3]/dl[3]/dd') %>% 
    html_text()

 # PARSE YEAR FOUNDED
  year_founded <- page %>% 
    html_nodes(xpath = '//*/section[1]/div[3]/dl[5]/dd') %>% 
    html_text()
  
  # PARSE EMPLOYEE COUNT
  employees <- page %>% 
    html_nodes(xpath = '//*/section[1]/div[3]/dl[6]/dd') %>% 
    html_text()

  # TEMP TO STORE LOOP DATA
  temp_df <- data.frame(page_url, revenue_millions, industry, year_founded, employees)
  
  # COMBINE TEMP WITH ALL DATA
  company_raw <- rbind(company_raw, temp_df) 
}
```

## Data cleaning
Ideally, we want to separate our data collection from our data processing but I did not anticipate "billions" to show up in the *parse revenue* step of our loop.

We will leave it as is though to illustrate the ease with which we can clean data in R, specifically the tidyverse.
```{r eval=FALSE}
# BRING BACK TO TIDYVERSE
company_data <- company_raw %>% 
  as_tibble() 

# EXTRACT 'BILLION' VALUES AND CONVERT TO MILLIONS
billions <- company_data %>% 
  filter(grepl(" Billion", revenue_millions)) %>% 
  mutate(revenue_millions = str_replace(revenue_millions, " Billion", ""),
         revenue_millions = as.numeric(as.character(revenue_millions)),
         revenue_millions = revenue_millions*1000)

# CONVERT 'MILLION' VALUES TO NUMERICAL FORMAT
company_data <- company_data %>% 
  filter(!grepl(" Billion", revenue_millions)) %>% 
  mutate(revenue_millions = as.numeric(as.character(revenue_millions)))

# JOIN OUR SANITIZED DATASET
company_data <- rbind(company_data, billions)
```

## Summarize data
With our cleaned dataset we can finally answer the question: *how much revenue did each industry generate in 2019 and how many companies were surveyed?*

```{r eval=FALSE}
company_parsed <- company_data %>% 
  group_by(industry) %>% 
  summarize(revenue_millions = sum(revenue_millions),
            count = n()) %>% 
  ungroup() %>% 
  filter(!is.na(revenue_millions)) %>% 
  mutate(pct_count = count / sum(count),
         pct_revenue = revenue_millions / sum(revenue_millions),
         revenue_billions = revenue_millions / 1000) 
```


```{r message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
library(kableExtra)

kable(top_20) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                font_size = 12, 
                full_width = F,
                fixed_thead = T)
```

Unsurprisingly, the *Health* industry comes out on top with a whopping $39Bn - nearly 1.7x more than the runner up *Consumer Products & Services*.

In terms of companies surveyed, we have *Business Products & Services* at #1 with 490 companies and a very close second for the *Advertising & Marketing* industry at 489.

### Keeping it real

<blockquote class="twitter-tweet" data-conversation="none" data-lang="en"><p lang="en" dir="ltr">Rather than a signal of the US economy, it&#39;s more a sign of which types of companies care about this (paid) PR exposure. Inc 5000 is a badge to show-off at agencies and software companies. Less so in other markets. It&#39;s not surprising that Health is the real winner (in the US) 🙂</p>&mdash; Rob Kerry (@robkerry) <a href="https://twitter.com/robkerry/status/1171684881031671809?ref_src=twsrc%5Etfw">September 11, 2019</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


## Visualize data
The tabulated data above is a little difficult to interpret so we'll plot our results instead.

```{r results='hide', message=FALSE, warning=FALSE, fig.width=10, fig.height=7}
library(ggrepel)

company_parsed %>% 
  ggplot(aes(pct_count, pct_revenue, label = industry)) +
  geom_point() +
  geom_label_repel() +
  geom_abline(color = 'salmon', lty = 'dashed') + 
  scale_x_continuous(labels = scales::percent_format(round(1)),
                     limits = c(0, .1)) +
  scale_y_continuous(labels = scales::percent_format(round(1)),
                     limits = c(0, .2)) +
  labs(x = "% of Total Companies", y = "% of Total Revenue",
       title = "Inc.5000 Fastest Growing Private Companies of 2019",
       subtitle = "Compares the number of companies surveyed to total revenue generated per industry",
       caption = "by:@eeysirhc\nsource:@viperchill") +
  theme_minimal()
```

We are pretty much *done* but there may be instances where we want data values to be accessible for users. Luckily, the [{plotly}](https://plot.ly/r/) package is our answer.

## Chart interactivity
```{r message=FALSE, warning=FALSE}
library(plotly)

plot_ly(data = company_parsed, 
        x = ~pct_count, 
        y = ~pct_revenue, 
        mode = "scatter",
        type = "scatter",
        size = 10,
        color = ~industry,
        colors = 'Set1', 
        hoverinfo = "text",
        text = ~paste("<b>Industry:</b> ",industry, 
                     "<br><b>Total Companies:</b> ", count,
                      "<br><b>Total Revenue ($Bn):</b> ", revenue_billions),
        showlegend = FALSE) %>% 
  layout(xaxis = list(title = "% of Total Companies",
                      tickformat = "%"),
         yaxis = list(title = "% of Total Revenue",
                      tickformat = "%"))
```


## Wrapping up
In my next article we'll take this a step further by building our own [Shiny app](https://shiny.rstudio.com). If you're feeling adventurous then you can use some [starter code here](https://www.christopheryee.org/blog/tidytuesday-housing-prices/) or checkout a barebones version for the [US housing price index](https://eeysirhc.shinyapps.io/20190205-housing_prices/).

As always, if you enjoyed this or found it helpful please share over your favorite internet medium!

