---
title: Visualizing Netflix viewing activity
author: Christopher Yee
date: '2019-07-02'
slug: visualizing-netflix-viewing-activity
categories:
  - Data Visualization
---

If you are like me then it's very likely you share your Netflix account with multiple users.

If you are also like me then it's very likely you would be curious about how your Netflix viewing activity coompares and contrasts to all the parasites on your account!

In this post we'll leverage [#rstats](https://twitter.com/search?q=%23rstats&src=typd) to visualize what that will look like.

## Load packages
Let's fire up our favorite packages.
```{r results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(igraph)
library(ggraph)
library(tidygraph)
library(influenceR)
```

## Download data
With the exception of my own viewing activity (I'm not ashamed!), I have provided anonymized Netflix viewing data from a few family and friends for you to follow along.
```{r results='hide', message=FALSE, warning=FALSE}
netflix_views_raw <- read_csv("https://raw.githubusercontent.com/Eeysirhc/netflix_views/master/netflix_views.csv")

netflix_views <- netflix_views_raw
```

## Peek at our raw dataset
We can draw a random sample from our dataset to see what we are working with.
```{r message=FALSE, warning=FALSE}
netflix_views %>% 
  sample_n(10)
```

A couple data processing steps immediately spring to mind:

* _Title_ is going to be messy if we do not canonicalize to the base name
* For the _User_ column we can replace the dummy values so it's easier to interpret later

## Transform the data

### Title - extract only the name of the show or movie
As always, we'll take a quick look at the data.
```{r message=FALSE, warning=FALSE}
netflix_views %>% 
  select(Title) %>% 
  tail() # ONLY LOOK AT THE LAST FEW ROWS OF DATA
```

We can then clean up our _Title_ variable with the **separate** function.
```{r message=FALSE, warning=FALSE}
netflix_views %>% 
  select(Title) %>% 
  tail() %>% 
  separate(Title, c("title"), sep = ":", extra ='drop') 
```

What we did with the above was drop everything after the first colon to remove season number along with its corresponding episode name.

### User - replace fake values with.....fake names
How do we know what we need to change? We can do a quick **count** on the _User_ column.
```{r message=FALSE, warning=FALSE}
netflix_views %>% 
  count(user, sort = TRUE)
```

The above result informs us we need to change four of those dummy values so we'll use a combination of **mutate** and **case_when** to achieve that.
```{r message=FALSE, warning=FALSE}
netflix_views %>% 
  sample_n(10) %>% 
  select(user) %>% 
  mutate(user = case_when(user == 'A_sibling' ~ "Apple",
                          user == 'B_sibling' ~ "Banana",
                          user == 'C_sibling' ~ "Cherry",
                          user == 'D_sibling' ~ "Dragon Fruit",
                          TRUE ~ user)) # THIS SAYS: IF NOTHING ELSE THEN USE THE ORIGINAL VALUE
```

Did it do what we wanted ?

```{r message=FALSE, warning=FALSE}
netflix_views %>% 
  #sample_n(10) %>% COMMENT THIS PORTION OUT SO WE DON'T TAKE A SAMPLE
  select(user) %>% 
  mutate(user = case_when(user == 'A_sibling' ~ "Apple",
                          user == 'B_sibling' ~ "Banana",
                          user == 'C_sibling' ~ "Cherry",
                          user == 'D_sibling' ~ "Dragon Fruit",
                          TRUE ~ user)) %>% 
  count(user, sort = TRUE)
```


### Putting it all together - [#tidyverse](https://twitter.com/search?q=%23tidyverse&src=typd) style

```{r message=FALSE, warning=FALSE}
netflix_clean <- netflix_views %>% 
  select(Date, user, Title) %>% # REARRANGE COLJUMNS
  separate(Title, c("title"), sep = ":", extra ='drop') %>% 
  mutate(user = case_when(user == 'A_sibling' ~ "Apple",
                          user == 'B_sibling' ~ "Banana",
                          user == 'C_sibling' ~ "Cherry",
                          user == 'D_sibling' ~ "Dragon Fruit",
                          TRUE ~ user)) %>% 
  mutate(Date = mdy(Date)) %>% 
  filter(Date >= Sys.Date() - 730)  # FILTER WITHIN THE PAST TWO YEARS
```

And let's validate our work:
```{r message=FALSE, warning=FALSE}
netflix_clean
```

Now that we are done cleaning our data we can finally build our network graph.

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">In Data Science, 80% of time spent prepare data, 20% of time spent complain about need for prepare data.</p>&mdash; Big Data Borat (@BigDataBorat) <a href="https://twitter.com/BigDataBorat/status/306596352991830016?ref_src=twsrc%5Etfw">February 27, 2013</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

## Build network graph
```{r message=FALSE, warning=FALSE}
netflix_graph <- netflix_clean %>% 
  count(user, title, sort = TRUE) %>% # COUNT THE NUMBER OF VIEWS PER USER BY TITLE
  filter(n >= 10) %>% # FILTER OUT LOW VOLUME DATA
  as_tbl_graph(directed = FALSE) %>% 
  mutate(group = group_infomap(),
         neighbors = centrality_degree(),
         center = node_is_center(),
         dist_to_center = node_distance_to(node_is_center()),
         keyplayer = node_is_keyplayer(k = 5)) %>% 
  activate(edges) %>% 
  filter(!edge_is_multiple()) %>% 
  mutate(centrality_e = centrality_edge_betweenness())
```

If you want more details on the above then I encourage you to check out this [amazing intro to the ggraph package](https://www.data-imaginist.com/2017/ggraph-introduction-edges/).

## Visualize network graph
```{r message=FALSE, warning=FALSE}
ggraph(netflix_graph, layout = 'kk') + 
  geom_edge_link(alpha = 0.5, color = 'grey54') +
  geom_node_point(aes(color = factor(group)), size = 3, show.legend = FALSE) +
  geom_node_text(aes(label = name), size = 3, repel = TRUE) +
  scale_color_brewer(palette = "Set1") +
  theme_void() +
  labs(caption = "by: @eeysirhc")
```

## Wrapping up
Now it's your turn - where do you lie in this network graph based on your own Netflix viewing activity? Do we watch similar TV shows or not really? How do you compare & contrast against your family and friends? What happens if you change the threshold from _10_ to _20_?

Don't hesitate to share your results on [Twitter](https://twitter.com/Eeysirhc)!


