---
title: Visualizing intraday SEM performance with R
author: Christopher Yee
date: '2019-12-23'
slug: visualizing-intraday-sem-performance-with-r
categories:
  - Data Visualization
---

Aside from the base bid, Google SEM campaign performance can be influenced by contextual signals from the customer. These include but are not limited to: device, location, gender, parental status, household income, etc.

For this post we'll focus on ad schedule (or intraday) and visualize how time of day and day of week is performing.

## Load data
```{r warning=FALSE, message=FALSE}
library(tidyverse)

# ANONYMIZED SAMPLE DATA
df <- read_csv("https://raw.githubusercontent.com/Eeysirhc/random_datasets/master/intraday_performance.csv")
```

### Spot check our data
```{r warning=FALSE, message=FALSE}
df %>% 
  sample_n(20)
```

## Clean data

### Convert to factors

The _day_of_week_ is a character and _time_of_day_ is a double data type. We need to transform them to factors so they don't surprise us later.

```{r warning=FALSE, message=FALSE}
df_clean <- df %>% 
  mutate(day_of_week = as.factor(day_of_week),
         hour_of_day = as.factor(hour_of_day)) 
```


### Reorder day of week

If we plot our chart now it won't be correct because _day_of_week_ will not be arranged properly from Sunday to Saturday format.
```{r warning=FALSE, message=FALSE}
levels(df_clean$day_of_week)
```

We can quickly refactor with a few lines of code:
```{r warning=FALSE, message=FALSE}
df_clean$day_of_week <- factor(df_clean$day_of_week, 
                               levels = c("Saturday", "Friday", "Thursday",
                                          "Wednesday", "Tuesday", "Monday", "Sunday"))
```


Let's check to make sure it did what we wanted:
```{r warning=FALSE, message=FALSE}
levels(df_clean$day_of_week)
```


## Plot our results
```{r warning=FALSE, message=FALSE}
df_clean %>% 
  ggplot(aes(hour_of_day, day_of_week, fill = conv_rate)) +
  geom_tile(color = 'white') +
  scale_fill_gradient2(low = 'grey', mid = 'white', high = 'steelblue') +
  labs(y = NULL, x = "Hour of Day", fill = "Conversion Rate") +
  theme_minimal() 
```

We can quickly see from this visualization that Tuesday & Wednesdays at midnight the conversion rate is 7% or higher. Conversely, 10AM on Saturday has a conversion rate close to the 0% range.

With this insight we can then amplify or dampen our intraday bid modifiers to improve overall campaign efficiencies.

### And for the love of data...

...[segment](https://www.christopheryee.org/blog/for-the-love-of-data-segment/)!

```{r wanring=FALSE, message=FALSE}
df_clean %>% 
  ggplot(aes(hour_of_day, day_of_week, fill = conv_rate)) +
  geom_tile(color = 'white') +
  scale_fill_gradient2(low = 'grey', mid = 'white', high = 'steelblue') +
  labs(y = NULL, x = "Hour of Day", fill = "Conversion Rate") +
  theme_minimal() +
  facet_grid(account ~ .)
```

