---
title: 'TidyTuesday: Steam Games'
author: Christopher Yee
date: '2019-07-30'
slug: tidytuesday-steam-games
categories:
  - TidyTuesday
---

Data from #tidytuesday week of 2019-07-30 ([source](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-07-30))

### Load R packages
```{r results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(RColorBrewer)
library(scales)
```

### Download data
```{r results='hide', message=FALSE, warning=FALSE}
steam_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-30/video_games.csv") 
```

### Parse data
```{r results='hide', message=FALSE, warning=FALSE}
steam_games <- steam_raw %>% 
  
  # VARIABLE FOR AGE OF GAME
  mutate(release_year = substring(release_date, 8, 12), # EXTRACT YEAR
         release_year = as.numeric(str_trim(release_year)), 
         release_year = case_when(release_year == 5 ~ 2015, # INCORRECT DATA POINT
                          TRUE ~ release_year),
         age = 2019 - release_year) %>% 
  
  # VARIABLE FOR MIN/MAX NUMBER OF OWNERS
  mutate(max_owners = str_trim(word(owners, 2, sep = "\\..")),
         max_owners = as.numeric(str_replace_all(max_owners, ",", "")),
         min_owners = str_trim(word(owners, 1, sep = "\\..")),
         min_owners = as.numeric(str_replace_all(min_owners, ",", ""))) %>% 
  
  # REMOVE VALUES WITH INCONSISTENT RELEASE_DATE FORMAT (n=37)
  filter(age < 15) %>% 
  
  # FILTER OUT STUDIO SOFTWARE
  filter(price < 150)
```

### Visualize data

Question: how many people still play games that are _X_ years old (on Steam) ?
```{r message=FALSE, warning=FALSE}
steam_games %>% 
  select(age) %>% 
  ggplot() +
  geom_histogram(aes(as.factor(age)), 
                 binwidth = 1, 
                 stat = 'count',
                 fill = 'steelblue',
                 alpha = 0.8) +
  scale_y_continuous(labels = comma_format()) +
  labs(x = "Years Since Release",
       y = NULL,
       title = "Count of active games on Steam since release",
       caption = "by: @eeysirhc\nsource: Steam Spy") +
  theme_light()
```

Question: What is the distrubtion of game ownership for the top publishers ?
```{r message=FALSE, warning=FALSE}
steam_games %>% 
  select(-number, -release_date) %>% 
  mutate(max_owners = as.factor(max_owners / 1000000)) %>% 
  group_by(publisher) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n >= 80, publisher != 'NA') %>%
  mutate(publisher = paste0(publisher, " (", n, ")"),
         publisher = reorder(publisher, desc(n))) %>% 
  ggplot(aes(publisher, max_owners, color = publisher)) +
  geom_jitter(show.legend = FALSE, size = 3, alpha = 0.5) +
  scale_color_brewer(palette = 'RdBu') +
  theme_light() +
  labs(x = NULL,
       y = "Estimated Game Ownership per Million",
       title = "Distribution of ownership for top game publishers",
       subtitle = "(n) indicates total titles released",
       caption = "by: @eeysirhc\nsource: Steam Spy") +
  theme(panel.background = element_rect(fill = 'grey95'),
        plot.background = element_rect(fill = 'grey95'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) 
```

Interesting takeaways

* __Mass Market__: whether it is their distribution network or brand name, publishers such as _Ubisoft (117)_ and _Square Enix (85)_ manage to consistently break into the 1M+ user base categories.
* __Some Hits Mostly Miss__: majority of the top publishers will fall in this category like _Strategy First (126)_ or _Degica (81)_ where a few game titles will break 50K or 100K users but majority will fall in the 20K bucket.
* __Niche Market__: publishers like _Big Fish Games (n=265)_ and _Choice of Games (n=92)_ have a lot of titles under their belts where each game acquires no more than 20K users. It is possible they do not want to compete against major publishers like Ubisoft or SEGA but instead choose to churn out games which cater to a super niche audience.
