---
title: 'Find your favorite Twitter user with the rtweet package'
author: Christopher Yee
date: '2019-08-25'
slug: find-your-favorite-twitter-user-with-the-rtweet-package
categories:
  - Miscellaneous
---

Do you know who your favorite person on Twitter is? Probably!

Did you ever want to quantify that statement? Probably not!

Are you curious to find out who someone else's favorite Twitter user is? Now you can with R!

The code below is brought to you by [Namita](https://twitter.com/nnstats) and her hilarious tweet:

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">face some possibly uncomfortable truths about yourself and others with 4 easy lines of code using <a href="https://twitter.com/hashtag/rtweet?src=hash&amp;ref_src=twsrc%5Etfw">#rtweet</a> and the <a href="https://twitter.com/hashtag/tidyverse?src=hash&amp;ref_src=twsrc%5Etfw">#tidyverse</a> <a href="https://t.co/JtRnzk0xu7">pic.twitter.com/JtRnzk0xu7</a></p>&mdash; Namita (@nnstats) <a href="https://twitter.com/nnstats/status/1165112790270578688?ref_src=twsrc%5Etfw">August 24, 2019</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
source("~/Desktop/credentials.R")
```

## Load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(rtweet)
```

## Get [Twitter API key](https://developer.twitter.com/en/docs/basics/apps/overview)
```{r eval=FALSE}
create_token(
  app = "app_name",
  consumer_key = "consumer_key",
  consumer_secret = "consumer_secret",
  access_token = "access_token",
  access_secret = "access_secret"
)
```

## Grab data
```{r}
twitter_handle <- "eeysirhc"

get_favorites(twitter_handle, n = 3000) %>% 
  group_by(screen_name) %>% 
  tally(sort = TRUE, name = '# of tweets liked') %>% 
  slice(1:10)
```


Future article: grabbing the top 10 favorite users of my top 10 favorite users and then creating a network graph to identify strong/weak relationships.

