---
title: 'TidyTuesday: Beer Production'
author: Christopher Yee
date: '2020-04-14'
slug: tidytuesday-beer-production
categories:
  - TidyTuesday
---

Data from [#tidytuesday](https://twitter.com/search?q=%23tidytuesday) week of 2020-03-31 ([source](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-31/readme.md))

## Load packages
```{r results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(gganimate)
library(gifski)
```

## Download data
```{r results='hide', message=FALSE, warning=FALSE}
beer_states_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv")
```

## Clean data
```{r results='hide', message=FALSE, warning=FALSE}
beer_total <- beer_states_raw %>% 
  
  # FILL NULL VALUES WITH 0
  replace(., is.na(.), 0) %>% 
  
  # REMOVE LINE ITEM FOR 'TOTAL'
  filter(state != 'total') %>% 
  
  # COMPUTE TOTAL BARRELS PER YEAR BY STATE  
  group_by(year, state) %>% 
  summarize(total_barrels = sum(barrels)) %>% 
  
  ungroup() 
```

## Create rankings
```{r results='hide', message=FALSE, warning=FALSE}
beer_final <- beer_total %>% 
  
  group_by(year) %>% 
  mutate(
    
    # CALCULATE RANKINGS BY TOTAL BARRELS PRODUCED EACH YEAR
    rank = min_rank(-total_barrels) * 1.0,
    
    # STATE TOTAL DIVIDE BY STATE RANKED #1 PER YEAR
    produced = total_barrels / total_barrels[rank == 1],
    
    # CLEANED TEXT LABEL
    produced_label = paste0(" ", round(total_barrels / 1e6, 2), " M")) %>% 
  
  group_by(state) %>% 
  
  # SELECT TOP 20
  filter(rank <= 20) %>% 
  
  ungroup()
```

## Animate bar chart
```{r message=FALSE, warning=FALSE}
p <- beer_final %>% 
  ggplot(aes(rank, produced, fill = state)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(rank, y = 0, label = state, hjust = 1.5)) +
  geom_text(aes(rank, y = produced, label = produced_label, hjust = 0)) +
  coord_flip() +
  scale_x_reverse() +
  theme_minimal(base_size = 15) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  transition_time(year) +
  labs(title = "US Beer Production by State",
       subtitle = "Barrels produced each year: {round(frame_time)}",
       caption = "by: @eeysirhc\nsource: Alcohol and Tobacco Tax and Trade Bureau",
       x = NULL, y = NULL) 

animate(p, nframes = 300, fps = 12,
        width = 1000, height = 800, renderer = gifski_renderer())
```

## Save .gif

Last but not least, don't forget to save the file to your working directory:
```{r eval=FALSE}
anim_save("toast.gif")
```

## Acknowledgements

* [Thomas Lin Pedersen](https://twitter.com/thomasp85) for the [{gganimate} package](https://github.com/thomasp85/gganimate)
* The contributors to this [StackOverflow answer](https://stackoverflow.com/questions/53162821/animated-sorted-bar-chart-with-bars-overtaking-each-other/53163549)


## Related

* [TidyTuesday: Milk Production](https://www.christopheryee.org/blog/tidytuesday-milk-production/)
* [TardyThursday: College Tuition, Diversity & Pay](https://www.christopheryee.org/blog/tardythursday-college-tuition-diversity-pay/)






