---
title: 'TidyTuesday: Cocktails'
author: Christopher Yee
date: '2020-05-26'
slug: tidytuesday-cocktails
categories:
  - TidyTuesday
editor_options: 
  chunk_output_type: console
---

Data from [#tidytuesday](https://twitter.com/hashtag/tidytuesday) week of 2020-05-26 ([source](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-26/readme.md))

> If you are looking for the R script then you can [find it here](https://github.com/Eeysirhc/tidytuesday/blob/master/20200526-cocktails/cocktails.R)

## Load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel)
library(FactoMineR)
```


## Download data
```{r message=FALSE, warning=FALSE}
bc_raw <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-26/boston_cocktails.csv') 
```

## Data processing

### Standardize cases
```{r message=FALSE, warning=FALSE}
bc_raw %>% 
  count(ingredient, sort = TRUE) %>% 
  filter(str_detect(ingredient, "red pepper sauce"))
```

Let's fix that by making all ingredient values to lower case:
```{r message=FALSE, warning=FALSE}
bc <- bc_raw %>% 
  mutate(ingredient = str_to_lower(ingredient)) %>% 
  distinct() %>% 
  select(name, ingredient)
```

And to make sure it works...
```{r message=FALSE, warning=FALSE}
bc %>% 
  count(ingredient, sort = TRUE) %>% 
  filter(str_detect(ingredient, "red pepper sauce"))
```


### Fix untidy data

To follow "tidy" principles, we need one row per observation. 
```{r message=FALSE, warning=FALSE}
bc %>% 
  filter(str_detect(ingredient, ",")) 
```

We can reformat this by separating the commas then adding a new row for each cocktail per ingredient.

```{r message=FALSE, warning=FALSE}
# CLEAN DATAFRAME
bc_tidy <- bc %>% 
  filter(!str_detect(ingredient, ","))

# EXTRACT UNTIDY DATA THEN CLEAN
bc_untidy <- bc %>% 
  filter(str_detect(ingredient, ",")) %>% 
  mutate(ingredient = str_split(ingredient, ", ")) %>% 
  unnest(ingredient)

# COMBINE BOTH DATAFRAMES
bc_clean <- rbind(bc_tidy, bc_untidy) %>% 
  distinct()

bc_untidy
```

### Reduce cardinality

Our dataset has more than 550 unique ingredients so let's trim that down to the ingredients that are used in ten or more cocktails.

```{r message=FALSE, warning=FALSE}
bc_clean %>% 
  distinct(ingredient) %>% 
  count()

n_ingredients <- bc_clean %>% 
  count(ingredient, sort = TRUE) %>% 
  filter(n > 10)
```


### Normalize ingredients

Similar to our case statement section above, let's make sure our ingredients are consolidated to the same format.

```{r message=FALSE, warning=FALSE}
df <- bc_clean %>% 
  inner_join(n_ingredients) %>% 
  select(-n) %>% 
  mutate(ingredient = str_replace_all(ingredient, "-", "_"),
         ingredient = str_replace_all(ingredient, " ", "_"),
         ingredient = str_replace_all(ingredient, "old_mr._boston_", ""),
         ingredient = str_replace_all(ingredient, "old_thompson_", "")) 
```

## Multiple Correspondence Analysis (MCA)

Our dataset is mostly categorical so [MCA](https://en.wikipedia.org/wiki/Multiple_correspondence_analysis) can help identify and highlight any underlying structures.

### Format data for MCA
```{r message=FALSE, warning=FALSE}
df_mca_processed <- df %>% 
  mutate(value = 1) %>%
  pivot_wider(names_from = ingredient) %>% 
  replace(is.na(.), 0) %>% 
  select(-name) %>%
  mutate_if(is.double, as.factor)

mca_results <- MCA(df_mca_processed, graph = FALSE)
```


### Shape data to tidy structure
```{r message=FALSE, warning=FALSE}
mca_df <- data.frame(mca_results$var$coord)

mca_final <- rownames_to_column(mca_df, var = "rowname") %>% 
  as_tibble() %>% 
  filter(str_detect(rowname, "_1")) %>% 
  mutate(variable = str_replace_all(rowname, "_1", "")) %>% 
  select(variable, everything(), -rowname) %>% 
  mutate(highlight = case_when(str_detect(variable, "gin") ~ "gin",
                               str_detect(variable, "rum") ~ "rum",
                               str_detect(variable, "vodka") ~ "vodka",
                               str_detect(variable, "whiskey") ~ "whiskey",
                               str_detect(variable, "brandy") ~ "brandy",
                               str_detect(variable, "bourbon") ~ "bourbon",
                               str_detect(variable, "tequila") ~ "tequila"))
```


### Final plot

```{r message=FALSE, warning=FALSE, fig.height=12, fig.width=20}
mca_final %>% 
  ggplot(aes(x = Dim.1, y = Dim.2, label = variable, color = highlight)) +
  geom_density2d(color = "gray90") +
  geom_point(show.legend = FALSE) +
  geom_text_repel(show.legend = FALSE) +
  labs(x = "D1", y = "D2", 
       title = "Multiple correspondence analysis (MCA) on the most common cocktail ingredients",
       subtitle = "Closer points suggest they are typically mixed together",
       caption  = "by: @eeysirhc\nsource: Mr. Boston Bartender's Guide") +
  theme_minimal(base_size = 15) +
  theme(axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) 
```



## Future work

* [Part 2](https://www.christopheryee.org/blog/tidytuesday-cocktails-pt-2/): cocktail recommendation system based on the input of favorite drink
* Calculate and use dissimilarity to recommend a drink to someone based on what they dislike

## References

* [Multiple correspondence analysis in R](https://rpubs.com/gaston/MCA)




