---
title: 'TidyTuesday: Cocktails pt.2'
author: Christopher Yee
date: '2020-05-28'
slug: tidytuesday-cocktails-pt-2
categories:
  - TidyTuesday
editor_options: 
  chunk_output_type: console
---

This is part 2 of [TidyTuesday: Cocktails](https://www.christopheryee.org/blog/tidytuesday-cocktails/).

Below shows how we can use [#rstats](https://twitter.com/hashtag/rstats) to write a cocktail recommendation system that takes in a drink and returns a few other cocktails based on similarly mixed ingredients.

## Load libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(recommenderlab)
```

## Download and parse data

> Note: please check out part 1 for deatils on processing steps

```{r message=FALSE, warning=FALSE}
bc_raw <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-26/boston_cocktails.csv')

bc <- bc_raw %>% 
  mutate(ingredient = str_to_lower(ingredient)) %>% 
  distinct() %>% 
  select(name, ingredient)

bc_tidy <- bc %>% 
  filter(!str_detect(ingredient, ","))

bc_untidy <- bc %>% 
  filter(str_detect(ingredient, ",")) %>% 
  mutate(ingredient = str_split(ingredient, ", ")) %>% 
  unnest(ingredient)

bc_clean <- rbind(bc_tidy, bc_untidy) %>% 
  distinct()

df <- bc_clean %>% 
  mutate(ingredient = str_replace_all(ingredient, "-", "_"),
         ingredient = str_replace_all(ingredient, " ", "_"),
         ingredient = str_replace_all(ingredient, "old_mr._boston_", ""),
         ingredient = str_replace_all(ingredient, "old_thompson_", "")) 

df_processed <- df %>% 
  mutate(value = 1) %>% 
  pivot_wider(names_from = name) %>% 
  replace(is.na(.), 0)
```

## Recommendation algorithm

### Transform data to binary rating matrix
```{r message=FALSE, warning=FALSE}
cocktails_matrix <- df_processed %>% 
  select(-ingredient) %>% 
  as.matrix() %>% 
  as("binaryRatingMatrix")
```

### Create evaluation scheme
```{r message=FALSE, warning=FALSE}
scheme <- cocktails_matrix %>% 
  evaluationScheme(method = "cross", k = 5, train = 0.8, given = -1)
```


### Input customer cocktail preference

Let's check the ingredients for a very simple cocktail:
```{r message=FALSE, warning=FALSE}
df %>% filter(name == 'Screwdriver')
```

Perfect. We can now use our recommendation model and predict what cocktail the customer might like based on similarly mixed ingredients.

```{r message=FALSE, warning=FALSE}
favorite_cocktail <- "Screwdriver"

recco_matrix <- df %>% 
  select(name) %>% 
  unique() %>% 
  mutate(value = as.numeric(name %in% favorite_cocktail)) %>%
  pivot_wider(names_from = name) %>% 
  replace(is.na(.), 0) %>% 
  as.matrix() %>% 
  as("binaryRatingMatrix")

recco <- Recommender(getData(scheme, 'train'), 
                      method = "IBCF",  
                      param = list(k = 5))

cocktails_predict <- predict(recco, newdata = recco_matrix, n = 5)
```

### Cocktail recommendation output
```{r message=FALSE, warning=FALSE}
cocktails_results <- as(cocktails_predict, 'list') %>% as_tibble()

cocktails_results
```

Finally, let's make sure our results have at least _vodka_ or _orange juice_ as their ingredients for the recommended cocktails:

```{r message=FALSE, warning=FALSE}
df %>% 
  filter(name %in% cocktails_results$`1`) 
```

Not bad at all!

## Out of scope

* Model evaluaiton
* Dynamic web app built in Shiny

## References

* [Market basket analysis with recommenderlab](https://towardsdatascience.com/market-basket-analysis-with-recommenderlab-5e8bdc0de236)




