---
title: R script for the CausalImpact package
author: Christopher Yee
date: '2020-05-19'
slug: r-script-for-the-causalimpact-package
categories:
  - Data Science
editor_options: 
  chunk_output_type: console
---

Google has an amazing [#rstats](https://twitter.com/search?q=%23rstats) package called [CausalImpact](https://google.github.io/CausalImpact/CausalImpact.html) to predict the counterfactual: what would have happened if an intervention did __not__ occur.

This is a quick technical post to get someone up and running rather than a review of its literature, usage, or idiosyncrasies

## Load libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(CausalImpact)
```

## Download (dummy) data
```{r message=FALSE, warning=FALSE}
df <- read_csv("https://raw.githubusercontent.com/Eeysirhc/random_datasets/master/cimpact_sample_data.csv")

df %>% sample_n(5)
```

## Shape data

Before we can run our analysis, the CausalImpact package requires three columns:

* Date (YYYY-MM-DD)
* Response/Treatment
* Control

If your data is already structured in the above format then feel free to skip to the next section.

Otherwise, we need to massage our (dummy) data frame from a long to wide format.

```{r message=FALSE, warning=FALSE}
df_clean <- df %>%
  dplyr::select(date, experiment_type, revenue) %>%
  pivot_wider(names_from = "experiment_type",
              values_from = "revenue") %>%
  dplyr::select(date, experiment, control)
```

And a quick spot check:

```{r message=FALSE, warning=FALSE}
df_clean %>% 
  arrange(date) %>% 
  head()
```

## Set parameters

The code below will:

* Set the intervention start date
* How many days forward/backward to compare from start date (I suggest full 7-day weeks)
* Construct appropriate date variables

```{r}
test_date <- as.Date("2020-04-23")

test_length <- 21

pre <- c(test_date-(test_length+1), test_date-1)
post <- c(test_date, (test_date+test_length))
```

Let's also make sure our date differences are correct:

```{r message=FALSE, warning=FALSE}
pre[2]-pre[1]
```

And the post period?
```{r message=FALSE, warning=FALSE}
post[2]-post[1]
```

Good to go!

## Run causal impact analysis
```{r message=FALSE, warning=FALSE}
df_impact <- CausalImpact(df_clean, pre, post)
```

## Plot results
```{r message=FALSE, warning=FALSE}
plot(df_impact)
```

## Analysis summary
```{r message=FALSE, warning=FALSE} 
summary(df_impact)
```


## Detailed analysis
```{r message=FALSE, warning=FALSE}
summary(df_impact, "report")
```



