---
title: How to interact with Slack from R
author: Christopher Yee
date: '2020-03-11'
slug: how-to-interact-with-slack-from-r
categories:
  - Data Science
---

I think my tweet speaks for itself:

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">Words can not express how excited I am to use this :D</p>&mdash; Christopher Yee (@Eeysirhc) <a href="https://twitter.com/Eeysirhc/status/1237388132800061440?ref_src=twsrc%5Etfw">March 10, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

The goal of this article is to document how to send [#rstats](https://twitter.com/hashtag/rstats) code and plots directly to Slack.

# Load packages
```{r eval=FALSE}
library(slackr)
library(slackteams)
library(slackreprex)
```

# Slack credentials

## Member ID

You can easily grab that from [this guide here](https://medium.com/@moshfeu/how-to-find-my-member-id-in-slack-workspace-d4bba942e38c).

## Slack Key ID

To retrieve your Slack key ID, [login here](https://slack.com/oauth/authorize?client_id=220157155520.220159943344&scope=incoming-webhook,files:read,files:write:user,chat:write:bot,chat:write:user,mpim:write,mpim:read,mpim:history,im:write,im:read,im:history,groups:write,groups:read,groups:history,channels:write,channels:read,channels:history,emoji:read,usergroups:read,users:read) and then follow the prompts.

## API token & webhook 

Last but not least, we can now use our newfound credentials and insert them into the following URL:

_https<!-- -->://slackr-auth.herokuapp.com/creds/**MEMBER_ID**/**SLACK_KEY_ID**_

The end result should provide the last pieces of our puzzle for the API token, webhook and the assigned channel for the Slack key ID.

## Authenticate

```{r eval=FALSE}
slackrSetup(channel = "#thinktank",
            incoming_webhook_url = "https://hooks.slack.com/services/0123456789",
            api_token = "ABCDEFGHIJKLMNOPQRSTUVWXYZ")
```

# Interacting from R

We're now ready for a [toast](https://www.youtube.com/watch?v=fq6R39wWXGM) test:

```{r eval=FALSE}
slackr("toast test from R")
```

Even though we selected a specific channel we can easily switch to a different one without going through the entire process again:

```{r eval=FALSE}
slackr("toast test from R", channel = "#warroom")
```

## Sending a plot

Let's try to recreate a plot I saw awhile back on [SEO](https://www.christopheryee.org/categories/search-engine-optimization/) twitter.

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)

technical <- tibble(website_size = seq(0, 100, 1),
                    importance = seq(0, 100, 1)) %>% 
  mutate(segment = "technical")

content <- tibble(website_size = rev(seq(0, 100, 1)),
                  importance = seq(0, 100, 1)) %>% 
  mutate(segment = "content")

df <- rbind(technical, content)

p <- df %>% 
  ggplot(aes(website_size, importance, color = segment)) +
  geom_point() +
  geom_line() +
  labs(color = NULL) +
  theme_light() +
  theme(legend.position = 'top')
```

Making sure the plot does what we want...
```{r}
p
```

...and then beam it up over to Slack:
```{r eval=FALSE}
ggslackr(p)
```

Too. Easy.

# Applications

Nearly 85% of my work communications (internal & client-facing) is completed in Slack so I can think of a few "real world" use cases for this:

* Automated daily, weekly, monthly reporting
* Incorporate analyses & insights then publish results directly to Slack
* Ping the appropriate channels when performance surpasses a statistical threshold
* Sharing code or final data output without jumping through hoops
