---
title: 'Crypto market cap: evaluating realistic price points'
author: Christopher Yee
date: '2021-08-13'
slug: crypto-market-cap-evaluating-realistic-price-points
categories:
  - Crypto
---

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
# LIBRARIES
library(tidyverse)
library(scales)

# SIMULATION FUNCTION
future_mkcap <- function(max_price, cap, price){
  df <- runif(1e3, 0.01, max_price) %>% 
    as_tibble() %>% 
    rename(future_price = value) %>% 
    mutate(mkcap = (future_price * cap) / price) 
  
  return(df)
}

# PLOTTING FUNCTION
future_cprice <- function(coin_name, coin_cap, coin_price, t1_price, t2_price){
  
  coin <- future_mkcap(t2_price*2, coin_cap, coin_price) 

t1_target <- coin %>% 
  filter(future_price >= t1_price) %>% arrange(future_price) %>% head(1)
t2_target <- coin %>% 
  filter(future_price >= t2_price) %>% arrange(future_price) %>% head(1)

coin %>% 
  ggplot(aes(future_price, mkcap)) +
  geom_line(size = 1) +
  
  ### CURRENT PRICE
  # VERTICAL
  geom_segment(x = coin_price, y = -Inf,
               xend = coin_price, yend = log10(coin_cap),
               lty = 2, color = 'grey55', size = 0.75) +
  # HORIZONTAL
  geom_segment(x = -Inf, y = log10(coin_cap),
               xend = coin_price, yend = log10(coin_cap),
               lty = 2, color = 'grey55', size = 0.75) +
  
  ### TIER 1 PRICE
  # VERTICAL
  geom_segment(x = t1_target$future_price, y = -Inf,
               xend = t1_target$future_price, yend = log10(t1_target$mkcap),
               lty = 2, color = 'steelblue', size = 0.75) +
  # HORIZONTAL
  geom_segment(x = -Inf, y = log10(t1_target$mkcap),
               xend = t1_target$future_price, yend = log10(t1_target$mkcap),
               lty = 2, color = 'steelblue', size = 0.75) +
  
  ### TIER 2 PRICE
  # VERTICAL
  geom_segment(x = t2_target$future_price, y = -Inf,
               xend = t2_target$future_price, yend = log10(t2_target$mkcap),
               lty = 2, color = 'red', size = 0.75) +
  # HORIZONTAL
  geom_segment(x = -Inf, y = log10(t2_target$mkcap),
               xend = t2_target$future_price, yend = log10(t2_target$mkcap),
               lty = 2, color = 'red', size = 0.75) +
  
  scale_x_continuous(labels = dollar_format(),
                     limits = c(0, t2_price*1.5)) +
  scale_y_log10(labels = dollar_format(round(2))) +
  labs(x = "Price", y = "Market Capitalization in Billions (log-scale)",
       title = paste0(coin_name, " (2021-08-13): $", round(coin_price,2), " at $", round(coin_cap,2), "B market cap"),
       subtitle = paste0("$", t1_price, ": with $", round(t1_target$mkcap,2), "B market cap", "\n", 
                         "$", t2_price, ": with $", round(t2_target$mkcap,2), "B market cap"),
       caption = paste0("by: @eeysirhc")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

}
```

It may be tempting for new cryptocurrency investors to buy "cheap" coins because they assume a price of \$0.03, for example, can turn them into future millionaires just by [HODL](https://www.urbandictionary.com/define.php?term=hodl)-ing until it hits \$100.

Unfortunately, this is a common misconception as that future price may never materialize. This perspective only considers the demand side of the equation and completely omits data about the supply side to realistically estimate future crypto prices.

This is where the concept of market capitalization can be valuable to understand.^[1]^ Here is one definition according to [Coinbase](https://www.coinbase.com/learn/crypto-basics/what-is-market-cap):

> _...is the total dollar value of all the shares of a company’s stock — or, in the case of Bitcoin or another cryptocurrency, of all the coins that have been mined. In crypto, market cap is calculated by multiplying the total number of coins that have been mined by the price of a single coin at any given time._

In other words an investor can use market cap to gauge the relative value between many different cryptocurrencies.

So, how is this useful in evaluating future prices for a specific crypto? If...

> __Market Cap = Current Price * Circulating Supply__ 

...then with a little math we can deduce the following:

> __Future Market Cap = (Desired Price * Current Market Cap) / Current Price__

With this formula we can plug in the variables and calculate for a particular cryptocurrency to achieve a desired \$X price point, based on existing market conditions it would need a \$Y future market cap...assuming all else remains equal.

That was a lot of information so let's use an example with [Bitcoin](https://coinmarketcap.com/currencies/bitcoin/). Back in January 1st, 2020 it had a closing price of \$7,200.17 and a market capitalization of \$130.581 billion. At the time, would you believe someone if they claimed Bitcoin would eventually surpass \$40K?

Regardless of your answer we can calculate what the expected market cap should be at that price:

> __725.433 billion = ($40,000 * $130.581 billion) / $7,200.17__

Thus, for Bitcoin to realize a price point of \$40K it would need over \$725 billion in market cap - more than 5x at the time. What was its market cap a year later when it finally broke \$40K on January 9th, 2021?

\$748.563 billion...off by -3%.

To make things interesting we can now extrapolate this and visualize the relationship between the desired price point and future market cap potential to see what is required for a \$70K or even \$100K Bitcoin price:

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
future_cprice("Bitcoin", 893379142227/1e9, 47557.84, 70000, 100000)
```

Want to see a Bitcoin price of \$100K? Then expect it to be valued at a \$1.9 trillion market cap which, quite frankly, may not be entirely unrealistic.

Let's switch gears and apply this formula to other cryptocurrencies.

[Cardano](https://coinmarketcap.com/currencies/cardano/), a popular altcoin, has price predictions all over the place. We'll use the two that I have seen quite frequently: __\$5__ and __\$50__.

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
future_cprice("Cardano", 64842202136/1e9, 2.02, 5, 50)
```

Which one do you think is more realistic for Cardano? 

__\$5__ where it needs to 2.5x its current market cap? Or __\$50__ where it not only needs to 25x but achieve \$1.6 trillion in market cap value...completely surpassing that of Bitcoin's current \$893 billion at the time of this writing?^[2]^

## Wrapping up

The future price of a "cheap" cryptocurrency may never materialize if someone does not factor in market capitalization value. With this simple formula I hope it is useful for judging how realistic a desired future price can be based on readily available financial indicators.

If you found this remotely helpful or entertaining then don't hesitate to drop some cryptocurrencies into my [cryptojar](https://www.christopheryee.org/cryptojar/). 

:)

<hr>

^[1]^ _The information provided here is for general educational purposes only and is not intended to constitute investment advice. Such information is not, and should not be read as, an offer or recommendation to buy or sell or a solicitation of an offer or recommendation to buy or sell any particular digital asset or to use any particular investment strategy._

^[2]^ _This is not claiming it won't or can't happen in the future but merely stating that chances are very slim based on existing market data._

<br>

