---
title: 'Recreating plots in R: Coast FIRE Grid'
author: Christopher Yee
date: '2025-06-10'
slug: recreating-plots-in-r-coast-fire-grid
categories:
  - Data Visualization
editor_options: 
  chunk_output_type: console
---

The tl;dr about [Coast FIRE](https://walletburst.com/coast-fire/)...

> Coast FIRE is when you have enough in your retirement accounts that without any additional contributions, your net worth will grow to support retirement at a traditional retirement age.

I encourage everyone to read up on this very interesting personal finance concept!

## Load packages
```{r message=FALSE, error=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
```

## CoastFIRE function
```{r message=FALSE, error=FALSE, warning=FALSE}
coastfire_calculator <- function(retirement_spend, 
                                 withdraw_rate, 
                                 growth_rate, 
                                 inflation_rate, 
                                 retire_age, 
                                 current_age){
  df <- (retirement_spend) / (withdraw_rate * (1+(growth_rate - inflation_rate))^(retire_age - current_age))
  df
}
```

## Formula parameters

Assumes 3.5% safe withdrawal rate, 8% growth, 3% inflation and retired at 67 years old.

```{r message=FALSE, error=FALSE, warning=FALSE}
coast_withdraw <- 0.035
coast_growth <- 0.08
coast_inflation <- 0.03
retire_age <- 67
```

## Generate dataset
```{r message=FALSE, error=FALSE, warning=FALSE}
coast_data <- expand.grid(retirement_spending = seq(20000, 100000, 5000),
            age = seq(20, 60, 2)) %>% 
  as_tibble() %>% 
  mutate(coast_fire = round(coastfire_calculator(retirement_spending, 
                                                 coast_withdraw, 
                                                 coast_growth, 
                                                 coast_inflation, 
                                                 retire_age, age) / 1000, digits = 0)) 
```

## Label data
```{r}
coast_final <- coast_data %>% 
  mutate(segment = case_when(coast_fire < 250 ~ '0 to $250k', 
                             coast_fire >= 250 & coast_fire < 500 ~ '$250k - $500k', 
                             coast_fire >= 500 & coast_fire < 750 ~ '$500k - $750k',
                             coast_fire >= 750 & coast_fire < 1000 ~ '$750k - $1M',
                             coast_fire >= 1000 & coast_fire < 1250 ~ '$1M - $1.25M', 
                             coast_fire >= 1250 & coast_fire < 1500 ~ '$1.25M - $1.5M', 
                             coast_fire >= 1500 ~ '$1.5M+')) %>% 
  mutate(segment = as.factor(segment))
```

## Rearrange factor levels
```{r}
coast_final$segment <- factor(coast_final$segment, 
                              levels = c("0 to $250k", "$250k - $500k", 
                                         "$500k - $750k", "$750k - $1M", 
                                         "$1M - $1.25M", "$1.25M - $1.5M", 
                                         "$1.5M+"))
```

## Visualize data
```{r message=FALSE, error=FALSE, warning=FALSE, fig.width=12, fig.height=10}
coast_final %>% 
  ggplot(aes(age, retirement_spending, fill = segment, label = coast_fire)) + 
  geom_tile(color = 'gray55') +
  geom_text(color = 'black') + 
  theme_bw() +
  scale_fill_manual(values = c("darkolivegreen3", "darkolivegreen2", "yellow1", 
                               "orange1", "darkorange1", "tomato2", "red3")) +
  scale_x_continuous(breaks = coast_final$age) + 
  scale_y_continuous(labels = dollar_format(),
                     breaks = coast_final$retirement_spending) + 
  labs(x = "Age", y = "Annual Spending in Retirement", fill = NULL) + 
  theme_bw(base_size = 15) + 
  theme(legend.position = 'none')

```

## Alternative

The original is a little more on the conservative side so let's modify slightly with 4% safe withdrawl rate, 10% growth, 3% inflation and retire at 62 instead. 

```{r message=FALSE, error=FALSE, warning=FALSE, fig.width=12, fig.height=10, echo=FALSE}
coast_withdraw <- 0.04
coast_growth <- 0.10
coast_inflation <- 0.03
retire_age <- 62

coast_data <- expand.grid(retirement_spending = seq(20000, 100000, 5000),
            age = seq(20, 60, 2)) %>% 
  as_tibble() %>% 
  mutate(coast_fire = round(coastfire_calculator(retirement_spending, 
                                                 coast_withdraw, 
                                                 coast_growth, 
                                                 coast_inflation, 
                                                 retire_age, age) / 1000, digits = 0)) 

coast_final <- coast_data %>% 
  mutate(segment = case_when(coast_fire < 250 ~ '0 to $250k', 
                             coast_fire >= 250 & coast_fire < 500 ~ '$250k - $500k', 
                             coast_fire >= 500 & coast_fire < 750 ~ '$500k - $750k',
                             coast_fire >= 750 & coast_fire < 1000 ~ '$750k - $1M',
                             coast_fire >= 1000 & coast_fire < 1250 ~ '$1M - $1.25M', 
                             coast_fire >= 1250 & coast_fire < 1500 ~ '$1.25M - $1.5M', 
                             coast_fire >= 1500 ~ '$1.5M+')) %>% 
  mutate(segment = as.factor(segment))

coast_final$segment <- factor(coast_final$segment, 
                              levels = c("0 to $250k", "$250k - $500k", 
                                         "$500k - $750k", "$750k - $1M", 
                                         "$1M - $1.25M", "$1.25M - $1.5M", 
                                         "$1.5M+"))

coast_final %>% 
  ggplot(aes(age, retirement_spending, fill = segment, label = coast_fire)) + 
  geom_tile(color = 'gray55') +
  geom_text(color = 'black') + 
  theme_bw() +
  scale_fill_manual(values = c("darkolivegreen3", "darkolivegreen2", "yellow1", 
                               "orange1", "darkorange1", "tomato2", "red3")) +
  scale_x_continuous(breaks = coast_final$age) + 
  scale_y_continuous(labels = dollar_format(),
                     breaks = coast_final$retirement_spending) + 
  labs(x = "Age", y = "Annual Spending in Retirement", fill = NULL) + 
  theme_bw(base_size = 15) + 
  theme(legend.position = 'none')
```

## Related

* [Recreating plots in R: the power of tiny gains](https://www.christopheryee.org/blog/recreating-plots-in-r-the-power-of-tiny-gains/)
* [Recreating plots in R: intro to bootstrapping](https://www.christopheryee.org/blog/recreating-plots-in-r-intro-to-bootstrapping/)

